---
title: "Word Clouds - Opioids"
author: "Alexander Turner"
output: html_document
---

## Required Packages

```{r warning=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(magrittr)
library(wordcloud2)
library(wordcloud)
```

## Word Cloud Packages

This project uses 2 R packages for word clouds:

* __[`wordcloud`](https://cran.r-project.org/web/packages/wordcloud/wordcloud.pdf)__
* __[`wordcloud2`](https://cran.r-project.org/web/packages/wordcloud2/vignettes/wordcloud.html)__

## Data 

The word clouds were based on a dataset containing [Reddit](https://www.reddit.com/) text from posts / comments about [opioids](https://www.drugabuse.gov/drugs-abuse/opioids). Find more info [here](https://www.kaggle.com/amalinow/18000-reddit-comments-about-opioids/home). This dataset was chosen as opioid abuse has become an [epidemic](https://www.hhs.gov/opioids/about-the-epidemic/index.html) in the USA and a word cloud of social media comments helps us understand the issue. 

To be clear, the dataset contains a number of different fields and this analysis focuses on the post titles. For example, one post is titled:

> "What is the best pain killer?"

A comment would then mention "opioid" for it to become part of the dataset. Some of titles mention "opioid" themselves. 

## Word Cloud using `wordcloud2`

Make sure to hover your mouse over each word to see the exact count with `wordcloud2`

```{r warning=FALSE, message=FALSE, fig.align='center'}
# read plain text file 
opioids_data <- read_delim("C:/RStudio Projects/Personal Projects/Word Cloud/reddit_opioids.txt",
                           delim = "|",
                           col_names = T)

# remove all non-letter characters, convert to lower case & split sentences
opioid_titles <- opioids_data$Title %>% 
  gsub("[^A-Za-z ]", "", .) %>% 
  tolower() %>% 
  strsplit(" ") %>% 
  unlist()

# calculate word frequencies & put in table
opioid_words <- opioid_titles %>% table()

# create data frame for word frequencies table & remove 1st row which is blank 
opioid_words <- data_frame(
  word = opioid_words %>% names(),
  freq = opioid_words %>% as.integer()
) %>% 
  slice(-1)

# filter out some very common uninteresting words & all words with a frequency of less than 5
opioid_words %>% 
  filter(
    freq >= 5
    & word != "the"
    & word != "to"
    & word != "and"
    & word != "a"
    & word != "of"
    & word != "is"
    & word != "in"
    & word != "on"
  ) %>% 
  wordcloud2(color = "red", fontFamily = "Consolas") # create word cloud
```

## Word Cloud using `wordcloud`

```{r warning=FALSE, message=FALSE, fig.align='center'}
# create data frame for second word cloud
opioid_words_filtered <- opioid_words %>% 
  filter(
    word != "the"
    & word != "to"
    & word != "and"
    & word != "a"
    & word != "of"
    & word != "is"
    & word != "in"
    & word != "on"
  )

# second word cloud
wordcloud(
  words = opioid_words_filtered$word,
  freq = opioid_words_filtered$freq,
  random.order = F,
  rot.per = 0.3,
  max.words = 275,
  colors = brewer.pal(5, "Set1")
  )
```
