---
title: "Word Clouds - Fake News"
author: "Alexander Turner"
output: html_document
---

## Required Packages

```{r warning=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(magrittr)
library(wordcloud2)
```

## Word Cloud Package

This project uses the __[`wordcloud2`](https://cran.r-project.org/web/packages/wordcloud2/vignettes/wordcloud.html)__ R package for generating word clouds

## Data 

The word cloud was based on a [fake news](https://en.wikipedia.org/wiki/Fake_news) text dataset. Find more info [here](https://www.kaggle.com/mrisdal/fake-news/home). This dataset was chosen as fake news plagued the 2016 USA election so much so that it was [suggested that it might have influenced the result](http://www.abc.net.au/news/2016-11-14/fake-news-would-have-influenced-us-election-experts-say/8024660). A word cloud quickly allows us to see which words were used most in fake news posts (the results are not surprising), providing insight into the topic.

To be clear, the dataset contains a number of different fields and this analysis focuses on the post titles. For example, one post is titled:

> "Assange - Hillary will win and tried to hack WikiLeaks"

## Word Cloud

Make sure to hover your mouse over each word to see the exact count with `wordcloud2`

```{r warning=FALSE, message=FALSE, fig.align='center'}
# read csv file
news_data <- read_csv("C:/RStudio Projects/Personal Projects/Word Cloud/fake.csv") %>% 
  filter(title != "")

# remove all non-letter characters, convert to lower case & split sentences
news_titles <- news_data$title %>% 
  gsub("[^A-Za-z ]", "", .) %>% 
  tolower() %>% 
  strsplit(" ") %>% 
  unlist()
  
# calculate word frequencies & put in table
news_words <- news_titles %>% table()

# create data frame for word frequencies table & remove 1st row which is blank  
news_words <- data_frame(
  word = news_words %>% names(),
  freq = news_words %>% as.integer()
) %>% 
  slice(-1)

# filter out some very common uninteresting words & all words with a frequency of less than 50
news_words %>% 
  filter(
    freq >= 50
    & word != "the"
    & word != "to"
    & word != "and"
    & word != "a"
    & word != "of"
    & word != "is"
    & word != "in"
    & word != "on"
    & word != "are"
    & word != "from"
    & word != "at"
    & word != "as"
    & word != "for"
  ) %>% 
  wordcloud2(color = "black", fontFamily = "Consolas") # create word cloud
```
